\chapter*{Acknowledgements}

\tolerance=1
\emergencystretch=\maxdimen
\hyphenpenalty=10000
\hbadness=10000

I would like to thank my supervisor, Professor Tore Selland Kleppe, for his constant support and invaluable guidance. You have allowed me to 
pursue ideas, to fail, and so many times steered me in the right direction with detailed feedback and questions that would uncover flaws, but which would eventually lead me closer towards truth.
%I feel very lucky having had such an inspirational figure as my supervisor.
Thank you for your encouraging words, enthusiasm and genuine thoughtfulness.
%, enthusiasm and encouraging words.
% the academic walks, and for all the coffee.

Thanks are also due to my co-supervisors, the professors Hans Julius Skaug and Jan Terje Kvaløy.
Professor Skaug has co-authored two of the papers in this thesis, allowed a stay at the University of Bergen, and always shared of his time and wide experience, for this I am grateful.
%, been a tremendous support allowed me to join on academic walks in Nygårdsparken and allowed a stay at the University of Bergen in the fall of 2017, for this I am grateful. 
%Academic walks in Nygårdsparken with someone that I truly admire is a memory I know to cherish.
% Discussing potential applications of ideas with someone I truly admire, while wandering in the beautiful Nygårdsparken is exactly the kind of memory I now cherish.
Professor Kvaløy has been presence of constant cheerfulness and inspiration. Thank you for always taking a genuine interest in people and their ideas.
I extend my thanks to my fellow PhD-students. Utmost appreciation goes out to Kjartan Kloster Osmundsen and Birthe Aarekol, for the many discussions and several trips all around the world to academic conferences and meetings.

% TAKK KJETIL!!!

Finally, I want to give many thanks to my friends, in particular Kjetil for his unlimited accommodating spirit. And also my family, in particular my mother, Katrin, for invaluable advice, and my wife, Saeron Min, which has been a force of continued support. 
%Without you I am likely to have crumbled under times of utmost pressure.
% this work would not have been possible.
%Under times of utmost pressure I have not crumbled because of you.
You have let me dive into hours of silent thoughts, calculations and coding when it was needed, but have also pulled me away and forced upon me a more balanced life when I would encounter a runtime error of the mind, but of the kind which I would not be able to see or solve by myself.
\vspace{60pt}

Berent Ånund Strømnes Lunde\\
Stavanger, August 2020